---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# mvtraits

The goal of `mvtraits` is to efficiently fit Bayesian (hierarchical) multivariate models to trait data.
A key feature is the ability to perform multiple imputation of partially missing data.

## Installation

This package is not currently available on CRAN.
To install directly from this repository, do the following:

```r
devtools::install_github("ashiklom/mvtraits")
```

## Example: The Iris dataset

This is a basic example showing how a multivariate distribution can be fit to the classic Iris dataset.

```{r example}
library(mvtraits)
data(iris)
```

### Multivariate fit

The input data must be a matrix, with each trait in its own column.
Below, we drop the `Species` column and convert the result to a matrix.

```{r matrix}
iris_mat <- as.matrix(iris[, !grepl("Species", colnames(iris))])
```


Fit a multivariate distribution with `fit_mvnorm`:

```{r fit}
iris_fit <- fit_mvnorm(iris_mat)
str(iris_fit, max.level = 2)
```

Compare the results to univariate summary statistics.

```{r summarystats}
iris_fit[[c("stats", "mu", "Mean")]]
colMeans(iris_mat)
iris_fit[[c("stats", "Sigma", "Mean")]]
cov(iris_mat)
```

### Partial missing data

Fitting models also works with partially missing data.
Here, we remove half of the data at random.

```{r add_missing}
iris_mat2 <- iris_mat
nmat <- length(iris_mat2)
iris_mat2[sample.int(nmat, floor(nmat / 2))] <- NA
head(iris_mat2)
```

Now, re-fit the model.

```{r fit_missing}
iris_fit2 <- fit_mvnorm(iris_mat2)
iris_fit2[[c("stats", "mu", "Mean")]]
iris_fit2[[c("stats", "Sigma", "Mean")]]
```

### Hierarchical model

`mvtraits` can also fit hierarchical models with the `fit_mvnorm_hier` function.
The arguments are the same as `fit_mvnorm`, but also need a `group` argument that takes a vector of groups (similar to base R's `tapply`).

```{r fit_hier}
iris_fit2h <- fit_mvnorm_hier(iris_mat2, iris[["Species"]])
str(iris_fit2h, max.level = 2)
```
